---
title: "Summer in Italy: A Seasonal Analysis of Airbnb Trends"
author: Luke Checchio
date: Fall 2024
output: 
  pdf_document:
    fig_caption: yes
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: load-packages

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tidymodels))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(MuMIn))
suppressPackageStartupMessages(library(patchwork))
suppressPackageStartupMessages(library(knitr))

opts_chunk$set(echo = F, warning = F, message = F, fig.show = "asis")
```

```{r}
#| label: load-data

city_grouped <- read.csv("data/airbnb_italian_city_grouped.csv") |>
  mutate(across(where(is.character), str_to_lower))
neighbourhood_grouped <- read.csv("data/airbnb_italian_neighbourhood_grouped.csv") |>
  mutate(across(where(is.character), str_to_lower))
city_data <- read.csv("data/airbnb_italy_city.csv") |>
  mutate(across(where(is.character), str_to_lower))
neighbourhoods_geo <- st_read("data/airbnb_italy_city_neighbourhoods_geojson.geojson", quiet = T) |>
  mutate(across(where(is.character), str_to_lower))
city_neighbourhoods <- read.csv("data/airbnb_italy_city_neighbourhoods.csv") |>
  mutate(across(where(is.character), str_to_lower))
```

# 1. Introduction

## 1.1 Motivation

After dropping to historic lows in the midst of the COVID-19 pandemic, travel is well on track to recover to pre-2020 levels. These trends show no sign of a slowing, and by 2030, international travel is expected to reach 9 billion lodging nights per year (McKinsey & Company).Since the advent of Airbnb, the landscape of short-term lodging and tourism has transformed dramatically, providing an unprecedented array of choices for travelers and a novel source of income for property owners. Italy, with its rich cultural heritage and popular tourist destinations, has seen significant growth in Airbnb listings, particularly in major cities and tourist hubs. In fact, today, Airbnb accounts for the second largest share of online bookings within the country, at a nearly 30% market share (Statista). The dynamics of pricing within the Italian Airbnb market, however, remain complex and subject to various influences, including seasonal fluctuations which significantly impact both availability and cost.

The interplay between seasonality and pricing is critical in tourism-centric economies like Italy's, where different regions experience peak tourist seasons at various times of the year. For example, coastal areas see a surge during summer months, while cities rich in art and history, such as Florence and Rome, typically can attract visitors year-round. Understanding these patterns is essential not only for economic forecasting and business planning for hosts but also for policy-making in tourism management.

This research project aims to delve into the seasonal trends in Airbnb pricing across Italy, exploring connections across 5 major Italian tourist hubs. By analyzing a comprehensive dataset of Airbnb listings that includes detailed pricing information, review scores, host details, and more from December 2023 to September 2024, this study seeks to uncover insights that could inform more strategic pricing and marketing decisions for Airbnb hosts and provide policymakers with data to better understand the economics of short-term rentals. The findings aim to contribute to the broader discourse on sustainable tourism in Italy, balancing economic benefits with community impacts and long-term viability.

## 1.2 Dataset and Variables

```{r}
#| label: data-modification-split

city_data <- city_data |>
  group_by(id) |>
  mutate(relative_price = price / mean(price), prop_reviews = number_of_reviews / sum(number_of_reviews), n = n())

city_data <- city_data |>
  group_by(place) |>
  mutate(relative_lat = abs(latitude - mean(latitude)), relative_long = abs(longitude - mean(longitude)))

city_data$log_price <- log(city_data$price)

firenze_data <- city_data |> filter(place == "firenze")
milano_data <- city_data |> filter(place == "milano")
napoli_data <- city_data |> filter(place == "napoli")
roma_data <- city_data |> filter(place == "roma")
venezia_data <- city_data |> filter(place == "venezia")
```

The dataset used for this study comprises Airbnb listing records from five major Italian cities: Florence, Milan, Naples, Rome, and Venice. Data were methodically scraped on a quarterly basis from December 2023 to September 2024 to capture seasonal variations in listing attributes, correlating each data collection round with one of the four seasons. This temporal alignment lends itself to a detailed examination of how seasonal factors influence Airbnb pricing, amongst other variables.

Data collection was conducted using a consistent unique identifier for each listing across all scraping sessions, facilitating longitudinal analysis of the same properties over time. Each record includes the listing price at the time of scraping, whether the host is designated as a "superhost" (a status awarded by Airbnb to hosts recognized for exceptional service), the neighborhood and precise coordinates of the listing, room type, maximum guest capacity, aggregate review scores, the city of the listing, and the season during which the data were collected. In total, the dataset encompasses 245,035 individual entries across 89,306 unique Airbnb listings.

Significant steps were undertaken to prepare and refine the dataset for analysis The price variable exhibited a right-skewed distribution with several extreme outliers; to address this, a logarithmic transformation was applied to normalize the data using the natural logarithm, significantly improving its suitability for regression analysis (details are elaborated in the Exploratory Analysis section and the Appendix). Additionally, the initial exploration revealed challenges in incorporating location data due to the high variability and large number of neighborhoods—up to 114 in Venice alone. To streamline the analysis and avoid the pitfalls of overfitting, standardized location metrics termed relative latitude and relative longitude were developed. These metrics measure each listing’s deviation from the average latitude and longitude of all listings within the same city, respectively, providing a standardized approach to quantify location effects within cities.

Additional datasets containing the same underlying data but grouped by city and neighbourhood as well as neighbourhood and city geospatial data for visualization purposes were used in exploratory analysis, but not ultimately included in the model fit.

## 1.3 Exploratory Analysis

Initial exploratory data analysis was conducted on a number of variables from the dataset to uncover potential trends in the data as well as to help guide variable selection for the eventual model.

```{r}
#| label: response-distribution
#| fig.cap: "Log-Transformed Airbnb Price Distribution"

ggplot(city_data, aes(x = log_price)) +
  geom_density(color = "#008C45", fill = "#CD212A", size = 1.2) +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#FFFFFF", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.title = element_text(face = "bold") 
  ) +
  labs(
    title = "Distribution of Log-transformed Airbnb prices in Italy",
    x = "Log of Price",
    y = "Density"
  )
```

The plot above shows the log-transformed distribution of Airbnb prices in Italy. The transformation normalizes the price data, making it easier to observe common price points and variability. A similar plot of the prices prior to log-transformation is included in the Appendix, where it is clear that the distribution exhibits a heavy right-skew.

```{r}
#| label: seasonal-price-rome
#| fig.cap: "Median Airbnb Price Distribution by Neighbourhood, Season (Rome)"

neighbourhoods_geo <- neighbourhoods_geo |>
  left_join(city_neighbourhoods, by = "neighbourhood") |>
  transmute(neighbourhood = neighbourhood, city = city.y, 
            neighbourhood_group = neighbourhood_group.x, geometry = geometry)

grouped_neighbourhoods_full <- neighbourhoods_geo |>
  right_join(neighbourhood_grouped, by = c("neighbourhood" = "neighbourhood", "city" = "place")) |>
  group_by(neighbourhood, city, period) |>
  mutate(period_price = mean(price_median)) |>
  ungroup()

grouped_neighbourhoods_full <- grouped_neighbourhoods_full |>
  mutate(period = case_when(period == "early spring" ~ "Spring",
                            period == "early autumn" ~ "Fall",
                            period == "early summer" ~ "Summer",
                            period == "early winter" ~ "Winter"))

grouped_neighbourhoods_full$period <- factor(grouped_neighbourhoods_full$period, 
                                             levels = c("Spring", "Summer", "Fall", "Winter"))

ggplot(grouped_neighbourhoods_full |> filter(city == "roma")) + 
  geom_sf(aes(fill = price_median)) +
  facet_grid(~ period) +
  scale_fill_gradient(low = "#f2d8d9", high = "#CD212A") +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#FFFFFF", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.title = element_text(face = "bold") ,
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(
    title = "Neighbourhood median Airbnb price by season, Rome",
    fill = "Median Price (Euro)"
  )
```

The above plot displays a series of adjacent maps illustrating the median Airbnb prices by neighborhood across different seasons in Rome. The maps use a color gradient from light pink to dark red to depict the range of median prices in euros, with darker shades indicating higher prices. From the visualization, it seems that across most neighbourhoods, the median price tends to be higher in Spring, Summer, and Fall than Winter. This relationship is especially prevalent within the center-most neighbourhood, the Historic Center (Centro Historico), which turns noticeably darker in the Summer and Fall, indicating higher median Airbnb prices. Graphs similar to that which is displayed above are included in the appendix for the four other cities examined. The relationship between season and price was less obvious in the plots for Napoli and Milan, while Florence and Venice's graphs both exhibited characteristics similar to that observed in Rome's.

# 2. Methodology

## 2.1 Model and Variable Selection

```{r}
#| label: model-fitting

city_lmer <- lmer(log_price ~ accommodates + review_scores_rating + room_type + host_is_superhost + period + relative_lat + 
                        relative_long + place + period*place + (1 | id), data = city_data)
firenze_lmer <- lmer(log_price ~ accommodates + review_scores_rating + room_type + host_is_superhost + period + relative_lat + 
                        relative_long + (1 | id), data = firenze_data)
milano_lmer <- lmer(log_price ~ accommodates + review_scores_rating + room_type + host_is_superhost + period + relative_lat + 
                        relative_long + (1 | id), data = milano_data)
napoli_lmer <- lmer(log_price ~ accommodates + review_scores_rating + room_type + host_is_superhost + period + relative_lat + 
                        relative_long + (1 | id), data = napoli_data)
roma_lmer <- lmer(log_price ~ accommodates + review_scores_rating + room_type + host_is_superhost + period + relative_lat + 
                        relative_long + (1 | id), data = roma_data)
venezia_lmer <- lmer(log_price ~ accommodates + review_scores_rating + room_type + host_is_superhost + period + relative_lat + 
                        relative_long + (1 | id), data = venezia_data)
```


These new measures allowed us to incorporate location relative to city center, which seemed highly relevant in the context of modeling Airbnb prices. In general, one might expect that the closer a listing is to the center of a city, the higher the price ought to be, due to 

## 2.2 Model Assumptions

# 3. Results

# 4. Discussion

\newpage

# Appendix

## References

https://www.mckinsey.com/industries/travel-logistics-and-infrastructure/our-insights/now-boarding-faces-places-and-trends-shaping-tourism-in-2024

https://www.statista.com/forecasts/1000846/hotel-private-accommodation-online-bookings-by-brand-in-italy

https://www.kaggle.com/datasets/salvatoremarcello/italian-airbnb-dataset?select=airbnb.csv

*Put citations here*

## Additional Exploratory Data Analysis

```{r}
#| label: response-distribution-original
#| fig.cap: "Airbnb Price Distribution"

ggplot(city_data, aes(x = price)) +
  geom_density(color = "#008C45", fill = "#CD212A", size = 1.2) +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#FFFFFF", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.title = element_text(face = "bold") 
  ) +
  xlim(10, 2500) +
  labs(
    title = "Distribution of Airbnb prices in Italy",
    x = "Price (Euros)",
    y = "Density",
    caption = "Note: Price observations greater than 2500 (n = 447) are excluded from this plot for clarity."
  )
```

```{r}
#| label: seasonal-price-florence
#| fig.cap: "Median Airbnb Price Distribution by Neighbourhood, Season (Florence)"

ggplot(grouped_neighbourhoods_full |> filter(city == "firenze")) + 
  geom_sf(aes(fill = price_median)) +
  facet_grid(~ period) +
  scale_fill_gradient(low = "#f2d8d9", high = "#CD212A") +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#FFFFFF", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.title = element_text(face = "bold") ,
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(
    title = "Neighbourhood median Airbnb price by season, Florence",
    fill = "Median Price (Euro)"
  )

```

```{r}
#| label: seasonal-price-milan
#| fig.cap: "Median Airbnb Price Distribution by Neighbourhood, Season (Milan)"

ggplot(grouped_neighbourhoods_full |> filter(city == "milano")) + 
  geom_sf(aes(fill = price_median)) +
  facet_grid(~ period) +
  scale_fill_gradient(low = "#f2d8d9", high = "#CD212A") +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#FFFFFF", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.title = element_text(face = "bold") ,
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(
    title = "Neighbourhood median Airbnb price by season, Milan",
    fill = "Median Price (Euro)"
  )

```

```{r}
#| label: seasonal-price-naples
#| fig.cap: "Median Airbnb Price Distribution by Neighbourhood, Season (Naples)"

ggplot(grouped_neighbourhoods_full |> filter(city == "napoli")) + 
  geom_sf(aes(fill = price_median)) +
  facet_grid(~ period) +
  scale_fill_gradient(low = "#f2d8d9", high = "#CD212A") +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#FFFFFF", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.title = element_text(face = "bold") ,
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(
    title = "Neighbourhood median Airbnb price by season, Naples",
    fill = "Median Price (Euro)"
  )

```

```{r}
#| label: seasonal-price-venice
#| fig.cap: "Median Airbnb Price Distribution by Neighbourhood, Season (Venice)"

ggplot(grouped_neighbourhoods_full |> filter(city == "venezia")) + 
  geom_sf(aes(fill = price_median)) +
  facet_grid(~ period) +
  scale_fill_gradient(low = "#f2d8d9", high = "#CD212A") +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#FFFFFF", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.title = element_text(face = "bold") ,
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(
    title = "Neighbourhood median Airbnb price by season, Venice",
    fill = "Median Price (Euro)"
  )
```


## Model Investigation

### Firenze Model

```{r}
#| label: model-checking-firenze

# model summary output
summary(firenze_lmer)

# model anova
anova(firenze_lmer)

# model marginal and conditional r-squared
r.squaredGLMM(firenze_lmer)

# model residual vs. fitted plot
plot(fitted(firenze_lmer), resid(firenze_lmer))

# model qq plot
qqnorm(resid(firenze_lmer))
qqline(resid(firenze_lmer))

# model residual density plot
plot(density(resid(firenze_lmer)))
```

### Milano Model

```{r}
#| label: model-checking-milano

# model summary output
summary(milano_lmer)

# model anova
anova(milano_lmer)

# model marginal and conditional r-squared
r.squaredGLMM(milano_lmer)

# model residual vs. fitted plot
plot(fitted(milano_lmer), resid(milano_lmer))

# model qq plot
qqnorm(resid(milano_lmer))
qqline(resid(milano_lmer))

# model residual density plot
plot(density(resid(milano_lmer)))
```

### Napoli Model

```{r}
#| label: model-checking-napoli

# model summary output
summary(napoli_lmer)

# model anova
anova(napoli_lmer)

# model marginal and conditional r-squared
r.squaredGLMM(napoli_lmer)

# model residual vs. fitted plot
plot(fitted(napoli_lmer), resid(napoli_lmer))

# model qq plot
qqnorm(resid(napoli_lmer))
qqline(resid(napoli_lmer))

# model residual density plot
plot(density(resid(napoli_lmer)))
```

### Roma Data

```{r}
#| label: model-checking-roma

# model summary output
summary(roma_lmer)

# model anova
anova(roma_lmer)

# model marginal and conditional r-squared
r.squaredGLMM(roma_lmer)

# model residual vs. fitted plot
plot(fitted(roma_lmer), resid(roma_lmer))

# model qq plot
qqnorm(resid(roma_lmer))
qqline(resid(roma_lmer))

# model residual density plot
plot(density(resid(roma_lmer)))
```

### Venezia Model

```{r}
#| label: model-checking-venezia

# model summary output
summary(venezia_lmer)

# model anova
anova(venezia_lmer)

# model marginal and conditional r-squared
r.squaredGLMM(venezia_lmer)

# model residual vs. fitted plot
plot(fitted(venezia_lmer), resid(venezia_lmer))

# model qq plot
qqnorm(resid(venezia_lmer))
qqline(resid(venezia_lmer))

# model residual density plot
plot(density(resid(venezia_lmer)))
```

### All Cities Model

```{r}
#| label: model-checking-all

# model summary output
summary(city_lmer)

# model anova
anova(city_lmer)

# model marginal and conditional r-squared
r.squaredGLMM(city_lmer)

# model residual vs. fitted plot
plot(fitted(city_lmer), resid(city_lmer))

# model qq plot
qqnorm(resid(city_lmer))
qqline(resid(city_lmer))

# model residual density plot
plot(density(resid(city_lmer)))
```

